Master Thesis

### 2013-05-23 my-work-rewritten.tex
Wrote 'Defining lexicons'

\begin{figure}
  \centering
  \begin{verbatim}
(ht->typer {{:head {:cat "n"}}      'N
            {:head {:cat "v"
                    :trans "true"}} (-> 'NP 'NP 'S)
            {:head {:cat "det"}}    (-> 'N (-> 'NP 'S) 'S)})
  \end{verbatim}
  \caption{\label{fig:lex-sig-impl} A lexicalized version of the
    $\Sigma$%_{Synt} TODO TODO TODO
    signature from~\ref{sssec:example-sig}.}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{verbatim}
(nonlex-sigr {'and?    (-> 'T 'T 'T)
              'imp?    (-> 'T 'T 'T)
              'forall? (-> (=> 'E 'T) 'T)
              'exists? (-> (=> 'E 'T) 'T)})
  \end{verbatim}
  \caption{\label{fig:nonlex-sig-impl} A signature of non-lexical
    constants belonging to $\Sigma$%_{Synt}
    from~\ref{sssec:acg-sem}. The single arrow \texttt{->} corresponds
    to linear implication while the double arrow \texttt{=>} corresponds
    to intuitionistic implication.}
\end{figure}

@@

>>>>

\begin{figure}
  \centering
\begin{verbatim}
(ht->typer {{:head {:cat "n"}}      'N
            {:head {:cat "v"
                    :trans "true"}} (-> 'NP 'NP 'S)
            {:head {:cat "det"}}    (-> 'N (-> 'NP 'S) 'S)})
\end{verbatim}
  \caption{\label{fig:lex-sig-impl} A lexicalized version of the
    $\Sigma$(\_{Synt}) signature from~\ref{sssec:example-sig}.}
\end{figure}

@@

\begin{figure}
  \centering
\begin{verbatim}
(nonlex-sigr {'and?    (-> 'T 'T 'T)
              'imp?    (-> 'T 'T 'T)
              'forall? (-> (=> 'E 'T) 'T)
              'exists? (-> (=> 'E 'T) 'T)})
\end{verbatim}
  \caption{\label{fig:nonlex-sig-impl} A signature of non-lexical
    constants belonging to $\Sigma$(\_{Sem})
    from~\ref{sssec:acg-sem}. The single arrow \texttt{->} corresponds
    to linear implication while the double arrow \texttt{=>} corresponds
    to intuitionistic implication.}
\end{figure}

@@

We have explained how we can use elements of our lexical database to
define signatures. Now, we will turn to lexicons. Similar to how we
reduced signatures from the formal triples $\mathopen{<}A, C,
\tau\mathclose{>}$ to just the type assignment function $\tau$, we can
similarly reduce lexicons from the pairs $\mathopen{<}F, G\mathclose{>}$
to only $G$, the part which operates on terms. $F$, or at least its
relevant subset, can be easily retrieved by putting the types of the
abstract constants side-by-side with the types of their object terms to
get a sufficient subset of $\hat{F}$ and finding $F$, the mapping from
atomic abstract types whose homomorphic extension includes $\hat{F}$.

In our relational framework, a mapping is most directly represented as a
binary relation, which is the representation that we have chosen. In the
previous subsubsection, we talked a lot about how individual entries in
the lexical database generate the items of a signature. Solving the
problem for signatures also solved it for lexicons, since their
structure is very similar: a signature is a set of pairs of constants
and types, a lexicon is a set of pairs of constants and
terms. Therefore, a lexical entry contributes the object terms for the
constants which it generated in the abstract signature.

Our representation of lexical constants was purposefully engineered so
as to make the lexical entry directly accessible to the lexicon
implementation. This way, the lexicon author can inspect the hypertag
and the wordform of the lexical entry, the specifier of the concrete
constant that it is mapping right now and even the type she assigned to
the constant in the abstract signature to allow her to choose the
correct object term. Furthermore, since lexicons are still
relations/sets, we can compose them together by disjunction. See
Figure~\ref{lex-impl} for the implementation of the lexicalized versions
of the $\mathcal{L}_{Syntax}$ and $\mathcal{L}_{Sem}$ lexicons.

\begin{figure}
  \centering
\begin{verbatim}
(defn syntax-lexo
  [synt-constant string-term]
  (with-sig-consts string-sig
    (l/fresh [hypertag string-constant]
      (has-hypertago synt-constant hypertag)
      (share-lex-entryo synt-constant string-constant)
      (string-sig string-constant)
      (fs-assigne hypertag string-term
                  {:head {:cat "n"}}
                  ,(rt string-constant)
                  {:head {:cat "v"
                          :trans "true"}}
                  ,(rt (ll [x y] (++ (++ x string-constant) y)))
                  {:head {:cat "det"}}
                  ,(ll [x R] (R (++ ua-stx-constant x)))))))

(defn sem-lexo
  [synt-constant sem-term]
  (with-sig-consts sem-sig
    (l/fresh [hypertag sem-constant]
      (has-hypertago synt-constant hypertag)
      (l/conde [(share-lex-entryo synt-constant sem-constant)
                (sem-sig sem-constant)
                (fs-assigne hypertag sem-term
                            {:head {:cat "n"}}
                            ,(rt sem-constant)
                            {:head {:cat "v"
                                    :trans "true"}}
                            ,(rt sem-constant))]
               [(fs-assigne hypertag sem-term
                            {:head {:cat "det"
                                    :det_type "indef"}}
                            ,(rt (ll [p q] (exists? (il [x] (and? (p x)
                                                                  (q x))))))
                            {:head {:cat "det"
                                    :det_type "def"}}
                            ,(rt (ll [p q] (forall? (il [x] (imp? (p x)
                                                                  (q x)))))))]))))

\end{verbatim}
  \caption{\label{fig:lex-impl} The implementation of the lexicons from
    subsection~\ref{ssec:acg}. NOTE: This is too verbose both to put on
    the paper and also to code. These examples demonstrates two common
    and related shapes of lexicons. The $\mathcal{L}$(\_{Syntax}) and
    the first half of the $\mathcal{L}$(\_{Sem}) are lexicalized, in the
    sense that the object term contains a lexical constant with the same
    lexical entry as the abstract constant.\footnote{This notion of
      ``lexicalized'' is quite different to the one I use in the rest of
      the thesis so far (which roughly means ``generated by a lexical
      database'') and also different from the one used by Kanazawa
      (FIX!).} This means that before constructing the object term, we
    have to acquire the object constant that is contained in the term
    and then we just fit it in of the recipes, which we choose according
    to the hypertag. The second half of $\mathcal{L}$(\_{Sem}) simply
    assigns object terms based on the hypertag. A more concise and
    desirable implementation would allow us simply write the pairs of
    hypertags and object terms (recipes encoded as lambdas). My plan is
    to have exactly this, but I have not thought yet about how to split
    the various aspects into different reusable abstractions.}
\end{figure}
